---
title: "Getting started with ivppmlhdfe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ivppmlhdfe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`ivppmlhdfe` fits instrumental-variable Poisson pseudo-maximum-likelihood
(IV-PPML) models with high-dimensional fixed effects. It is an R translation of
the Stata command of the same name (Kwon, Larch, Yoon and Yotov, 2026), and
reproduces `ppmlhdfe` / `fixest::fepois` point estimates to machine precision in
the plain (no-instrument) case.

> **Attribution.** This package is *only a translation*. All credit for the
> estimator and methodology goes to the original authors, Ohyun Kwon, Mario
> Larch, Jangsu Yoon and Yoto V. Yotov. The R translation is by Guewen Heslan
> (PhD Candidate in Economics, Nantes Université, France) using Claude Opus,
> available at <https://github.com/gheslan/ivppmlhdfe-for-r>.

```{r}
library(ivppmlhdfe)
data(ivppmlhdfe_ClassA)
```

## The endogeneity problem

In the Monte-Carlo design the endogenous regressor `x1` has a true coefficient
of `0.5`, but it is correlated with the error, so ordinary PPML is biased:

```{r}
ppml <- ivppmlhdfe(y ~ x1 + x2 | id + year, data = ivppmlhdfe_ClassA)
coef(ppml)["x1"]   # biased upward, away from 0.5
```

Instrumenting `x1` with `z` removes the bias:

```{r}
iv <- ivppmlhdfe(y ~ x2 | id + year | x1 ~ z, data = ivppmlhdfe_ClassA)
coef(iv)["x1"]     # close to the true 0.5
summary(iv)
```

## Two- and three-way gravity models

```{r}
data(ivppmlhdfe_ClassB)
b <- ivppmlhdfe(y ~ x2 | exp^year + imp^year | x1 ~ z,
                data = ivppmlhdfe_ClassB, cluster = ~pair)
coef(b)["x1"]
```

`^` denotes an interacted fixed effect (Stata's `#`). Standard errors can be
heteroskedasticity-robust (the default) or one-/two-way cluster-robust via the
`cluster` argument.

## Split-panel jackknife bias correction

Replacing the regressor with an instrument re-introduces incidental-parameter
bias. The split-panel jackknife removes its leading-order term. For Class A the
estimator splits along both the time and the cross-sectional dimension:

```{r}
s <- ivppml_spj(y ~ x2 | id + year | x1 ~ z, data = ivppmlhdfe_ClassA,
                class = "A", id = "id", time = "year", B = 0)
s$b_full   # uncorrected
s$b_spj    # SPJ-corrected
```

Set `B > 0` (the paper uses `B = 1000`) to obtain pair/individual-bootstrap
standard errors and a percentile confidence interval.

## Validating against ppmlhdfe / fepois

```{r}
if (requireNamespace("fixest", quietly = TRUE)) {
  fe <- fixest::fepois(y ~ x1 + x2 | id + year, data = ivppmlhdfe_ClassA,
                       notes = FALSE)
  rbind(ivppmlhdfe = coef(ppml)[c("x1", "x2")],
        fepois     = coef(fe)[c("x1", "x2")])
}
```

The two sets of coefficients agree to machine precision.

## References

Kwon, O., M. Larch, J. Yoon, and Y.V. Yotov (2026). *Instrumental-Variable
Poisson PML with High-Dimensional Fixed Effects.* CESifo Working Paper 12641.
